---
title: "Funded US+Canada Health Agency Grants"
author: "Shamsi Soltani"
date: "2023-07-01"
output: html_document
---
This doc unifies code from SICCS Summer 2023 in Montreal for my own purposes. I pull US NIH data and Canadian CIHR data and filter for suicide-related research.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)
library(pacman)
p_load(knitr, haven,
       tidyverse, tidytext, textdata, tm, 
       maps, SnowballC, wordcloud, topicmodels,
       repoRter.nih,magrittr,rvest, cld2)

p_load(stopwords,quanteda,quanteda.textmodels,quanteda.textstats,quanteda.textplots)
    #more text manipulation packages
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

#1.0 NIH Data
##1.1 prep headers for NIH API
```{r}
#define fields you want 
data("nih_fields") 

fields <- nih_fields %>%
  filter(response_name %in% 
    c("appl_id", "terms", "activity_code", "project_title", "fiscal_year", "award_amount", "is_active", "project_start_date", "abstract_text","program_officers", "activity_type", "contact_pi_name", "full_foa" , "organization", "organization_type", "agency_ic_fundings"
      )) %>% pull(include_name)
```

##1.2 create json for NIH API
```{r json, echo=FALSE}
#create json for NIH API
req <- make_req(criteria = list(advanced_text_search 
                    = list(operator = "advanced",
                    search_field = c("terms"
                                  #   , "abstract", "projecttitle"
                                  ),
                    search_text = "(suicide) OR \"self-directed violence\""),
            # activity_codes = 'F31',
             fiscal_years = 2000:2023
             ),
      include_fields = fields )
```
##1.3 retrieve data
```{r}
#grants <- get_nih_data(query = req, max_pages = 1, #flatten_result=TRUE)

#remove max pages once query final if seeking more than 500 recs; 
grants <- get_nih_data(query = req,  flatten_result=TRUE)
  #Note: API request fails after 15,000 records
```
##1.4 choose grant types & filter
```{r NIH grants types}
#activity codes ---> F, K grants for students, early career faculty (excludes institutional grants)
awards = c('F05', 'F30', 'F31', 'F32', 'F33', 
'F37', 'F38', 'F99', 'FI2', 'FM1',
#'R00',  'R01', 'R03', 'R13', 'R15', 'R16', 'R18', 
# 'R21','R24', 'R25', 'R28', 'R2F', 'R30', 
# 'R33', 'R34', 'R35', 'R36',  'R37', 'R38', 'R41', 'R42', 
# 'R43', 'R44', 'R49', 'R50',  'R55', 'R56', 'R61', 'R90', 
# 'RC1', 'RC2', 'RC3', 'RC4',  'RF1', 'RL1', 'RL2', 'RL5', 
# 'RL9', 'RM1', 'RS1', 
'K00', 'K01', 'K02', 'K05', 'K06', 
'K07', 'K08', 'K12', 'K14', 
'K18', 'K21', 'K22', 'K23', 
'K24', 'K25', 'K26', 'K30', 'K38', 'K43', 
'K76', 'K99', 'KD1', 'KL1', 'KL2', 'KM1' )
```

```{r filter}
table(grants$activity_code) 
  #N=225 F31 funded apps in last decade with suicide or self-directed violence as a title, narrative, or key word; 340 total in database

#filter out grant types we don't want
grants_filt = grants %>% 
  filter(activity_code %in% awards)
```

#2.0 CIHR data
##2.1 pull csv data
```{r, echo=FALSE}
setwd("/Users/ssoltani/Desktop/PhD coursework/Cours_SICSS_Montreal/SICSS_Grant-Scraping/00.Data/CIHR")

# store all csv file names
csv_grants <- list.files(getwd(), pattern = "*.csv")

# load csv files and merge them into one dataframe
cihr_data <- data.frame() 

for (file in csv_grants) {
  file_path <- file
  csv_data <- read.csv(file_path, check.names = F)
  cihr_data <- bind_rows(cihr_data, csv_data)
}

# Data filtering/restructuring 
names(cihr_data)
#removes accents from var
names(cihr_data)<-iconv(names(cihr_data), to = "ASCII", sub = "")
```

```{r}
cihr_grants <- cihr_data %>% 
  select('Key-Cl', 'Name-Nom', OrgNm, OrgID, ProvinceEN, CountryEN,
         'CompetitionYear-AnneDuConcours', 'FiscalYear-AnneFinancire',
         'AwardAmountinFY-MontantSubventiondeAF', 'TotalAwardAmount-MontantSubventionTotal',
         ProgramIDProgramme, ProgramNameEN, FundingTypeEN, AreasOfScienceEN, ResearchClassEN,
         TitreApplicationTitle, 'Keywords-MotsCls', 'Abstract-Rsum') %>% 
  rename(abstract = 'Abstract-Rsum',
         key = 'Key-Cl',
         name = 'Name-Nom',
         organization = OrgNm,
         orgID = OrgID,
         province = ProvinceEN,
         country = CountryEN,
         year_comp = 'CompetitionYear-AnneDuConcours', 
         year_fisc = 'FiscalYear-AnneFinancire',
         award_amount_FY = 'AwardAmountinFY-MontantSubventiondeAF',
         award_amount_total = 'TotalAwardAmount-MontantSubventionTotal',
         programID = ProgramIDProgramme,
         program_name = ProgramNameEN,
         funding_type = FundingTypeEN,
         science_areas = AreasOfScienceEN,
         research_class = ResearchClassEN,
         title = TitreApplicationTitle,
         keywords = 'Keywords-MotsCls') %>% 
  mutate(abstract = ifelse(nchar(iconv(abstract, to = "ASCII", sub = "")) <= 30, NA, abstract))
```

##2.2 Clean province column
```{r}
#cleaning province 
prov_terr= c('Alberta', 'Ontario', 'British Columbia', 'Québec', 'Qu<e9>bec','Nova Scotia', 'New Brunswick','Manitoba', 'Prince Edward Island', 'Saskatchewan', 'Newfoundland and Labrador','North West Territories','Nunavut', 'Yukon')

cihr_grants = cihr_grants %>% 
  mutate(province=ifelse(province %in% c('Unknown/Inconnu', 'Unknown'), 'Unknown', province),
         province=ifelse(!province %in% c(prov_terr, 'Unknown'), 'Hors du Canada', province),
         province=ifelse(province %in% c('Québec', 'Qu<e9>bec'), 'Quebec', province)) 

prov= cihr_grants%>% 
  group_by(province) %>% 
  summarise(ct=n()) 

ggplot(prov) + geom_col(aes(y=reorder(province,ct), x= ct)) +
    ggtitle("nombre des award/grant par province, toutes années")
```
#3.0 Combine data
joining US and CA data
```{r}
ca_data=suicide_topic
us_data=grants_filt

#names(us_data)

ca_data = suicide_topic %>% 
  select(year_fisc, funding_type, award_amount_FY, province, organization, keywords, abstract, title, name) %>% 
  mutate(country="CA", year_fisc=substr(year_fisc, 1, 4))
  
us_data = us_data %>% select(fiscal_year, activity_code, award_amount, organization_org_state, organization_org_name, terms, abstract_text, project_title, contact_pi_name) %>% 
arrange(fiscal_year, activity_code, award_amount, organization_org_state, organization_org_name, terms, abstract_text, project_title, contact_pi_name) %>% 
mutate(country="US")

names(us_data)= names(ca_data)
suicide_data=rbind(ca_data, us_data)
```

##3.1 adding language into 2-country dataset using title and abstracts' predicted lang
```{r}
#req cld2 pkg
suicide_data= suicide_data %>% 
  mutate(language_abstr = detect_language(abstract),
         language_title = detect_language(title))

table(suicide_data$country)

#crosstabs
suicide_data %>%
  group_by(country, language_abstr, language_title) %>%
  tally() %>%
  spread(country, n)

#test = suicide_data %>%
#  filter(language_abstr!=language_title)
    #the one discordant record is fr lang; note abstract predicted correctly (more text)

suicide_data= suicide_data %>% 
  mutate(lang = ifelse((language_abstr=='fr'|language_title=='fr'),'fr','en'),
         lang = ifelse((is.na(lang) & (language_abstr=='en'|language_title=='en')),'en',lang)) %>% 
  select(-c(language_abstr, language_title))
```